\chapter{Artefacts}

Artefacts produced during this projects are saved in the following repositories.

\section{Project Private Repository}

URL: https://svn.ecdf.ed.ac.uk/repo/user/s1024484/

This svn repository includes 


\begin{itemize}
  \item annual reports from 2010/2011 to 2012/2013
  \item a copy of the TAkka repository, with the source code of Erlang, Akka, 
and TAkka implementations of the the ATM example \citep{atmprivate} and the 
elevator controller example \citep{quviq}.
  \item a copy of the TAkka Play Repository.
  \item a copy of the FreeBench Repository.
%  \item a copy of the Rejuvenation Repository.
  \item an extension of the Links programming language that supports 
join-patterns (reference).
  \item a copy of this thesis and a [paper] summaries this project.
\end{itemize}


\section{TAkka Public repository}

URL: https://github.com/Jiansen/TAkka

This github repository includes the source code of the TAkka library, testing 
examples and instructions of using the code.  The ATM example \citep{atmprivate} 
and the elevator controller example \citep{quviq} are not included in this 
public repository for copyright reasons.  The TAkka implementation for the Play 
framework is in a separate repository.

\section{TAkka Play}
URL: https://github.com/Jiansen/TAkka-Play

This github repository includes the source code of the TAkka 
implementation of the Play framework.


\section{FreeBench Repository}
URL: https://github.com/Jiansen/FreeBench

This github repository includes the source code of the FreeBench tool, 
described in Section \ref{throughput}.


\begin{comment}
\section{Rejuvenation Repository}

URL: https://github.com/Jiansen/Rejuvenation


This github repository includes the source code of the TAkka library, testing 
examples except the ATM example \citep{atmprivate} and the elevator controller 
example \citep{quviq}, and scripts for running scalability tests on Beowulf 
cluster.

\end{comment}