\documentclass[preprint, authoryear, 10 pt]{sigplanconf}

\setpagenumber{1}

\usepackage{comment}

%% \usepackage[style=alphabetic]{biblatex}
%% \usepackage[authoryear]{natbib}
%% \PassOptionsToPackage{authoryear}{natbib}
%% \usepackage{caption}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\usepackage{paralist}
\usepackage[pdfpagelabels]{hyperref}
\usepackage[all]{hypcap}
\usepackage{verbatim}
\usepackage{array}
\usepackage{float}
\usepackage{multirow}
% \usepackage{rotating}
\usepackage{multicol}
\usepackage{longtable}

\usepackage{listings} 	
% "define" Scala
\lstdefinelanguage{scala}{
  morekeywords={abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@, <:<},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\usepackage{pxfonts}
\usepackage{color}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
 
% Default settings for code listings
\lstset{frame=tb,
  language=scala,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left, % none,
  numbersep=2pt,
  numberstyle=\tiny\color{gray},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  frame=none, % single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=2,
  captionpos=b
}

% A comment in a draft (shouldn't appear in the final version).
\newcommand{\mycomment}[1]{\(\spadesuit\){\bf #1 }\(\spadesuit\)}
% Comment this out in the draft
% \newcommand{\mycomment}[1]{}
\newcommand{\pmcomment}[1]{\comment{PM}{#1}}
\newcommand{\pwtcomment}[1]{\comment{PT}{#1}}
\newcommand{\rscomment}[1]{\comment{RS}{#1}}

% line break in table.  e.g.
% Foo bar & \specialcell{Foo\\bar} & Foo bar \\    % vertically centered
% Foo bar & \specialcell[t]{Foo\\bar} & Foo bar \\ % aligned with top rule
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\permission {[Copyright notice will appear here once ’preprint’ option is removed.]}

\begin{document}

% \markboth{V. F. Pamplona et al.}{Photorealistic Models for Pupil Light Reflex and Iridal Pattern Deformation}

\title{Type-parameterized Actors and Their Supervision (v6.6.5) } % title

\authorinfo{Jiansen HE}
           {University of Edinburgh}
           {jiansen.he@ed.ac.uk}
\authorinfo{Philip Wadler}
           {University of Edinburgh}
           {wadler@inf.ed.ac.uk}
\authorinfo{Philip Trinder}
           {University of Glasgow}
           {P.W.Trinder@glasgow.ac.uk}

% \category{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}[Animation]
% \category{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling}[Physically based modeling]

% \terms{Experimentation, Human Factors}



\maketitle



\begin{abstract} 
The robustness of distributed message passing applications can be improved by
\begin{inparaenum}[(i)]
 \item employing failure recovery mechanisms such as the supervision principle, or
 \item using typed messages to prevent ill-typed communication.
\end{inparaenum}
The former approach is originally developed in the Erlang OTP (Open Telecom 
Platform) library.  The later approach has been well explored in systems 
including the join-calculus and the typed $\pi$-calculus.  Attempts of combing 
those two approaches has been made in two directions: type checking existing 
Erlang programs and implementing a statically typed actor library that supports 
the supervision principle.   Unfortunately, current statically typed actor 
libraries either use dynamically typed messages (e.g. Akka), or do not support 
supervision (e.g. Cloud Haskell).

Implementing a statically typed actor library that also supports the 
supervision principle raises three problems. Firstly, distributed resources are 
dynamically typed whereas local implementation is statically type checked and 
does not need to consider ill-typed messages.  Therefore, a mixture of statical 
and dynamical type checking is required to make sure that distributed resources 
are well-typed.  Secondly, preventing unexpected messages to an actor requires 
that actor having different types when it communicates with others.  Thirdly, 
an elegant notion of sub-typing is required for the supervision purpose so that 
a supervisor can interact with child actors of different types, even in systems 
where sub-typing is not directly supported.




\begin{comment}
Implementing a statically typed actor library that also supports the 
supervision principle raises three problems.  Firstly, current design of name 
servers are dynamically typed and do not map typed names to resources of 
corresponding types.  Therefore, a novel typed name server is required.  
Secondly, actors cannot be naively parameterized by the union type of all 
expected message types, including those for supervision purposes, because 
supervisors must interact with child actors of different types.  Thirdly, 
unexpected messages can be sent to actors which publish more type information 
than necessary to other parties.
\end{comment}

This paper introduces the typed Akka library, TAkka, which resolves above
problems. Although TAkka actor extends Akka actor and use the Scala 
{\tt Manifest} class to serialize type information, we believe that similar 
improvements 
can be made to actor libraries in other languages.

We evaluate the TAkka library by re-implementing examples built from small and 
medium sized Erlang and Akka libraries.  Results show that Akka programs can be gradually upgraded 
to TAkka equivalents with minimal runtime and code size overheads.  
TAkka programs have similar scalability to their Akka equivalents. 
Finally, we port the Chaos Monkey library for assessing application reliability 
and design a Supervision View library for dynamically capturing the structure 
of supervision trees.

\begin{comment}
This paper presents the design, implementation and preliminary evaluation
results of the TAkka library, where types of actor related components are
statically typed and operations are type checked at the earliest possibility.
We introduce actors parameterized by the type of messages they expect to
receive. We show that it is straightforward to construct
supervision trees of type-parameterized actors. We minimize the number of system
messages that may be handled by general users by providing standard supervision
strategies.


\end{comment}


\end{abstract}

\category{D.1.3}{Programming Techniques}{Concurrent Programming}

\terms
Design, Languages, Reliability

\keywords
actor, type, supervision tree, name server


\input{intro}
\input{background}
\input{typechecking}
\input{actor}
\input{evolution}
\input{evaluation}
\input{reliability}


\section{Conclusion}

Existing actor libraries accept dynamically typed messages.  The TAkka
library introduces a type-parameter for actor-related classes. The additional 
type-parameter of a TAkka actor specifies the communication interface of 
that actor.  With the help of type-parameterized actors, unexpected 
messages to actors are rejected at compile time.

In addition to eliminating programming bugs and type errors, 
programmers would like to have a failure recovery mechanism for
unexpected run-time errors.  We are glad to see that type-parameterized 
actors can form supervision trees in the same way as untyped actors.

Lastly, test results show that building type-parameterized actors on top of 
Akka does not introduce significant overheads, with respect to program size,
efficiency, and scalability.  In addition, debugging techniques such 
as Chaos Monkey and Supervision View can be applied to applications built 
using actors with supervision trees.  The above results encourage the use of 
types and supervision trees to implement reliable applications and improve the 
reliability of legacy applications with little effort.  We expect similar 
results can be obtained in other actor libraries such as future extensions 
of CloudHaskell \cite{OTPCloudHaskell}.

\acks
Acknowledgments


\bibliographystyle{abbrvnat}
\bibliography{takka}

\appendix
%  \input{appendix_correct}

% \received{September 2008}{March 2009}

\end{document}


