\section{Introduction}

The Actor model defined by \citet{Hewitt:1973} treats actors as
primitive computational components.  Actors collaborate by sending asynchronous
messages to each other.  An actor independently determines its reaction to
messages it receives.  The Actor model is adopted by the Erlang programming 
language \citep{ArmstrongErlang} for building real time telephony applications. 
Erlang developers later propose the supervision principle \citep{OTP}, which 
suggests that actors should be organized in a tree structure so that any failed 
actor can be properly restarted by its supervisor.  Nevertheless, adopting the 
Supervision principle is optional in Erlang.

The notions of actors and supervision trees have been ported to statically 
typed languages including Scala and Haskell.  Scala actor libraries including 
Scala Actors \citep{actor_1, actor_2} and Akka \citep{akka_api,akka_doc} use 
dynamically typed messages even though Scala is a statically typed language.  
Some recent actor libraries, including Cloud Haskell \citep{CloudHaskell}, Lift 
\citep{lift_scala} , and scalaz \citep{scalaz}, support both 
dynamically and statically typed messages, but do not support supervision.
Can actors in supervision trees be statically typed?

The key claim in this paper is that actors in supervision trees can be
typed by parameterizing the actor class with the type of messages it expects to
receive.  Type-parameterized actors benefit both users and developers
of actor-based services. For users, sending ill-typed messages is prevented at 
compile time.  Because messages are usually transmitted asynchronously, it may 
be otherwise difficult to trace the source of errors at runtime, especially in 
distributed environments.  For service developers, since unexpected messages 
are eliminated from the system, they can focus on the logic of the services 
rather than worrying about incoming messages of unexpected types.  

\begin{comment}
Implementing 
type-parameterized actors in a statically-typed language; however, requires 
solving following three problems.

\begin{enumerate}
  \item A typed name server is required to retrieve actor references of
specific types.  A distributed system usually requires a name server that
maps names of services to processes that implement that service.  If processes
are dynamically typed, this is usually implemented as a map from names to 
processes. Can this be adapted to cases where processes are statically 
typed?

  \item Supervisor actors must interact with child actors of different types.  
Actors are structured in supervision trees to improve system reliability.  Each 
actor in a supervision tree needs to handle messages within its specific 
interests and also messages from its supervisor.  Is it practical to define a 
supervisor that communicates with children of different type parameters?

  \item Actors which receive messages from distinct parties may suffer from 
the type pollution problem, in which case a party imports too much type  
information about an actor and can send the actor messages not expected from it.
Systems built on layered architecture or the MVC model are often victims of the 
type pollution problem. As an actor receives messages from distinct parties 
using its sole channel, its type parameter is the union type of all expected 
message types.  Therefore, unexpected messages can be sent to actors which 
naively publishes its type parameter or permits dynamically typed messages. 
Can a type-parameterised actor have different types of communication interface, 
i.e. typed actor reference, when published to parties.
\end{enumerate}


\end{comment}

Continuing a line of work on merging types with actor programming by Haller
and Odersky \citep{actor_1, actor_2} and Akka developers \citep{akka_doc}, 
along with the work on system reliability test by 
\citet{ChaosMonkey} and \citet{ErlangChaosMonkey}, this paper makes 
following contributions.

\begin{itemize}
 \item It presents the design and implementation of a typed name server
that maps typed names to values of the corresponding type, for example actor 
references.  The typed name server (Section \ref{nameserver}) mixes 
static and dynamic type checking so that type  errors are detected at the earliest 
opportunity.  The implementation requires language support for type reflection 
and first class type descriptors.  In Scala, the {\tt Manifest} class provides 
such facility.

 \item It describes the design of the TAkka library.  Sections \ref{actor} to
\ref{actor_context} illustrate how type parameters are added to actor related
classes to improve type safety.   By separating the handler for system 
messages from the handler for user defined messages, Sections \ref{supervision}
and \ref{systemmessage} show that type-parameterized actors can form
supervision trees in the same manner as untyped actors.  Section 
\ref{evolution} explains how to upgrading Akka programs to their TAkka 
equivalents gradually.
 
 \item It gives a comprehensive evaluation of the TAkka library.  Section 
\ref{type_pollution} shows that the TAkka library can avoid the type pollution 
problem straightforwardly.  Results in
Section \ref{expressiveness} confirm that using type parameterized actors 
sacrifice neither expressiveness nor correctness.  Efficiency and 
scalability test in Section \ref{efficiency} shows that TAkka applications have 
little overhead at the initialization stage but have almost identical run-time 
performance and scalability compared to their Akka equivalents.  In 
addition, two helper libraries explained in Section \ref{reliability} are 
shipped with the TAkka library for assisting reliability evaluation.  The 
first library, Chaos Monkey, is ported to test applications against intensive 
adverse conditions.  The second library, Supervision View, is 
designed and implemented to dynamically monitor changes of supervision tree.
 
\end{itemize}
