% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{comment}



%% \usepackage[style=alphabetic]{biblatex}
\usepackage[authoryear,sectionbib]{natbib}
%% \PassOptionsToPackage{authoryear}{natbib}
%% \usepackage{caption}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\usepackage{paralist}
\usepackage[pdfpagelabels]{hyperref}
\usepackage[all]{hypcap}
\usepackage{verbatim}
\usepackage{array}
\usepackage{float}
\usepackage{multirow}
% \usepackage{rotating}
\usepackage{multicol}
\usepackage{longtable}

\usepackage{listings} 	
% "define" Scala
\lstdefinelanguage{scala}{
  morekeywords={abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@, <:<},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}


\usepackage{pxfonts}
\usepackage{color}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
 
% Default settings for code listings
\lstset{frame=tb,
  language=scala,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left, % none,
  numbersep=2pt,
  numberstyle=\tiny\color{gray},
  keywordstyle=\bfseries\color{black},
  commentstyle=\color{dkgreen},
  stringstyle=\color{black},
  frame=none, % single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=2,
  captionpos=b
}

% A comment in a draft (shouldn't appear in the final version).
\newcommand{\mycomment}[1]{\(\spadesuit\){\bf #1 }\(\spadesuit\)}
% Comment this out in the draft
% \newcommand{\mycomment}[1]{}
\newcommand{\pmcomment}[1]{\comment{PM}{#1}}
\newcommand{\pwtcomment}[1]{\comment{PT}{#1}}
\newcommand{\rscomment}[1]{\comment{RS}{#1}}

% line break in table.  e.g.
% Foo bar & \specialcell{Foo\\bar} & Foo bar \\    % vertically centered
% Foo bar & \specialcell[t]{Foo\\bar} & Foo bar \\ % aligned with top rule
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
  
  
\renewcommand\bibname{References}


\begin{document}
%
\frontmatter          % for the preliminaries
\title{Type-parameterized Actors and Their Supervision 
% (v7.0.1)
}

% \author{Jianse HE\inst{1} \and Philip Wadler\inst{1}
%  \and
% Philip Trinder \inst{2}}



%\authorrunning{J. HE P. Wadler P. Trinder} % abbreviated author list (for
%running head)

%

%%%% list of authors for the TOC (use if author list has to be modified)

% \tocauthor{Jianse HE, Philip Wadler, and Philip Trinder}

%

% \institute{University of Edinburgh \and
% University of Glasgow}


\maketitle



\begin{abstract} 
The robustness of actor-based concurrent applications can be Improved by (i) 
employing failure recovery mechanisms such as the supervision principle, or (ii) 
using typed messages to prevent ill-typed communication. This paper introduces 
TAkka, a typed Akka library that supports both typed messages and the 
supervision principle. The TAkka library mixes statical and dynamical type 
checking to make sure that dynamically typed distributed resources and 
statically typed local resources have consistent types. Our notion of typed 
actor can publish itself as different types when used by different parties so 
that messages of unexpected types are prevented at the senders' side. In TAkka, 
messages for supervision purposes are treated specially so that a supervisor can 
interact with child actors of different types. Results show that Akka programs 
can be gradually upgraded to TAkka equivalents with minimal runtime and code 
size overheads. Finally, we implement two auxiliary libraries for 
reliability assessment.
\end{abstract}

\begin{comment}
The robustness of distributed message passing applications can be improved by
\begin{inparaenum}[(i)]
 \item employing failure recovery mechanisms such as the supervision 
principle in the Erlang/OTP library, or
 \item using typed messages to prevent ill-typed communication.
\end{inparaenum}
Attempts of combing these two approaches have been made in two directions: type checking existing 
programs written in dynamically typed languages and implementing a statically 
typed actor library that supports the supervision principle.   Regarding the 
later approach, current statically typed actor libraries either use dynamically 
typed messages (e.g. Akka), or do not support supervision (e.g. Cloud Haskell 
and others).
This paper introduces the typed Akka library, TAkka, which supports both typed messages
and the supervision principle.  The TAkka library mixes statical and and dynamical type
checking to make sure that dynamically typed distributed resources and statically
typed local resources have consistent types.  Our notion of typed actor can 
publish itself as different types when used by different parties so that
messages of unexpected types are prevented at the sender's side.  In TAkka,
messages for supervision purposes are treated specially so that a supervisor can
interact with child actors of different types.
Results show that Akka programs can be gradually upgraded to TAkka equivalents 
with minimal runtime and code size overheads.   Finally, we port the Chaos 
Monkey library and design a Supervision View library to assist reliability 
evaluation.
\end{comment}





\input{intro}
\input{background}
\input{typechecking}
\input{actor}
\input{evolution}
\input{evaluation}
\input{reliability}


\section{Conclusion}

Existing actor libraries accept dynamically typed messages.  The TAkka
library introduces a type-parameter for actor-related classes. The additional 
type-parameter of a TAkka actor specifies the communication interface of 
that actor.  With the help of type-parameterized actors, unexpected 
messages to actors are rejected at compile time.
In addition to eliminating programming bugs and type errors, 
programmers would like to have a failure recovery mechanism for
unexpected run-time errors.  We are glad to see that type-parameterized 
actors can form supervision trees in the same way as untyped actors.
Lastly, test results show that building type-parameterized actors on top of 
Akka does not introduce significant overheads, with respect to program size,
efficiency, and scalability.  In addition, debugging techniques such 
as Chaos Monkey and Supervision View can be applied to applications built 
using actors with supervision trees.  The above results encourage the use of 
types and supervision trees to implement reliable applications and improve the 
reliability of legacy applications with little effort.  We expect similar 
results can be obtained in other actor libraries such as future extensions 
of CloudHaskell \cite{OTPCloudHaskell}.

\paragraph{Acknowledgements}


The authors gratefully acknowledge the substantial help they have received from 
many colleagues who have shared their related results and ideas with us over 
the long period during which this paper was in preparation.  
Benedict Kavanagh and Danel Ahman for continuous comments and discussions.
The RELEASE team for giving us the access to the source code of the BenchErl 
benchmark examples.  Thomas Arts from Quivq.com and Francesco Cesarini from 
Erlang Solutions for providing the Erlang source code for the ATM simulator 
example and the Elevator Controller example, two examples used in their 
commercial training courses.

% ??


%
% ---- Bibliography ----
\bibliographystyle{abbrvnat}
\bibliography{takka}

\end{document}
