% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass[authoryear,preprint]{sigplanconf}


%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{comment}



%% \usepackage[style=alphabetic]{biblatex}
% \usepackage[authoryear,sectionbib]{natbib}
%% \PassOptionsToPackage{authoryear}{natbib}
%% \usepackage{caption}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\usepackage{paralist}
\usepackage{verbatim}
\usepackage{array}
\usepackage{float}
\usepackage{multirow}
% \usepackage{rotating}
\usepackage{multicol}
\usepackage{longtable}

\usepackage{listings} 	
% "define" Scala
\lstdefinelanguage{scala}{
  morekeywords={abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@, <:<},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}


\usepackage{pxfonts}
\usepackage{color}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
 
% Default settings for code listings
\lstset{frame=tb,
  language=scala,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left, % none,
  numbersep=2pt,
  numberstyle=\tiny\color{gray},
  keywordstyle=\bfseries\color{black},
  commentstyle=\color{dkgreen},
  stringstyle=\color{black},
  frame=none, % single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=2,
  captionpos=b
}

\usepackage[pdfpagelabels]{hyperref}
\usepackage[all]{hypcap}


% A comment in a draft (shouldn't appear in the final version).
\newcommand{\mycomment}[1]{\(\spadesuit\){\bf #1 }\(\spadesuit\)}
% Comment this out in the draft
% \newcommand{\mycomment}[1]{}
\newcommand{\pmcomment}[1]{\comment{PM}{#1}}
\newcommand{\pwtcomment}[1]{\comment{PT}{#1}}
\newcommand{\rscomment}[1]{\comment{RS}{#1}}

% line break in table.  e.g.
% Foo bar & \specialcell{Foo\\bar} & Foo bar \\    % vertically centered
% Foo bar & \specialcell[t]{Foo\\bar} & Foo bar \\ % aligned with top rule
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
  
  
\renewcommand\bibname{References}


\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}
           
\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Typecasting Actors: from Akka to TAkka
}
% \subtitle{Subtitle Text, if any}

\authorinfo{Jiansen HE}
           {University of Edinburgh}
           {jiansen.he@ed.ac.uk}
\authorinfo{Philip Wadler}
           {University of Edinburgh}
           {wadler@inf.ed.ac.uk}
\authorinfo{Philip Trinder}
           {University of Glasgow}
           {Phil.Trinder@glasgow.ac.uk}

\maketitle



\begin{comment} 

Scala supports actors and message passing with the Akka library. Though Scala is 
statically typed, messages in Akka are dynamically typed (that is, of type Any). 
The Akka designers argue that using static types is "impossible" because "actor 
behaviour is dynamic", and, indeed, it is not clear that important actor 
support, such as supervision or name servers, could be implemented if messages 
are statically typed. Here we present Takka, a variant of Akka where messages 
are statically typed, and show that it is possible to implement supervisors and 
name servers in such a framework. We show it is possible to upgrade from Akka to 
Takka gradually, porting one module at a time; and we show that Takka can 
support behavioral upgrade where the new message type of an actor is a supertype 
of the old type. We demonstrate the expressiveness of Takka by rewriting a score 
of Akka applications; the percentage of lines that needs to be changed varies 
from 44\% (in a 25 line program) to 0.05\% (in a 27,000 line program), with a 
geometric mean of 8.5\%. We measure the execution speed, scalability, and 
throughput of Takka programs and show that it is comparable to Akka. [perhaps 
give precise details]

\end{comment}




\input{mainbody}
% \input{mainbody2}

% \input{intro}
% \input{background}
% \input{typechecking}
% \input{actor}
% \input{evolution}
% \input{evaluation}
% \input{reliability}




\acks


The authors gratefully acknowledge the substantial help they have received from 
many colleagues who have shared their related results and ideas with us over 
the long period during which this paper was in preparation.  
Benedict Kavanagh and Danel Ahman for continuous comments and discussions.
The RELEASE team for giving us access to the source code of the BenchErl 
benchmark examples.  Thomas Arts from Quviq.com and Francesco Cesarini from 
Erlang Solutions for providing the Erlang source code of two examples used in their 
commercial training courses.

\vspace{-10pt}


% ---- Bibliography ----
\bibliographystyle{abbrvnat}
\bibliography{takka}

\end{document}