% \documentclass[10pt, journal, compsoc, authoryear]{IEEEtran}
\documentclass[10pt, journal, compsoc, authoryear]{article}

% \setpagenumber{1}g
\usepackage{CJKutf8}


\usepackage{comment}

\usepackage[numbers]{natbib}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\usepackage{amsthm}
\usepackage{amsmath}
 \numberwithin{equation}{subsection}
\usepackage{paralist}
\usepackage[pdfpagelabels]{hyperref}
\usepackage[all]{hypcap}
\usepackage{verbatim}
\usepackage{array}
\usepackage{float}
\usepackage{multirow}
% \usepackage{rotating}
\usepackage{multicol}
\usepackage{longtable}
\usepackage{graphicx}


\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{conj}{Conjecture}
%\newtheorem{exmp}{Example}[section]
\newtheorem{exmp}{Example}
\newenvironment{where}{\noindent{}where\begin{itemize}}{\end{itemize}}
% A comment in a draft (shouldn't appear in the final version).
\newcommand{\mycomment}[1]{\(\spadesuit\){\bf #1 }\(\spadesuit\)}
% Comment this out in the draft
% \newcommand{\mycomment}[1]{}
\newcommand{\pmcomment}[1]{\comment{PM}{#1}}
\newcommand{\pwtcomment}[1]{\comment{PT}{#1}}
\newcommand{\rscomment}[1]{\comment{RS}{#1}}

% line break in table.  e.g.
% Foo bar & \specialcell{Foo\\bar} & Foo bar \\    % vertically centered
% Foo bar & \specialcell[t]{Foo\\bar} & Foo bar \\ % aligned with top rule
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

% \permission {[Copyright notice will appear here once ’preprint’ option is 
% removed.]}

\begin{document}


\title{Towards Precise Analysis for Software Rejuvenation \\ (v 0.1) } % title

\author{Jiansen HE}

% \category{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}[Animation]
% \category{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling}[Physically based modeling]

% \terms{Experimentation, Human Factors}



\maketitle


\begin{comment} 
\begin{abstract} 

Software rejuvenation refers to a preventive failure recovery technique, which 
restarts a software regularly to reduce the cost of software ageing.  It has been  
actively studied in the literature and applied in enterprise level applications. 
Traditionally, a software system is modelled as a Semi-Markov process, from 
which a rejuvenation schedule is deducted to optimise the objective function 
(e.g. highest availability, lowest cost) after the system reaches its steady-state.
Unfortunately, it can takes a long time for a system to reach its steady-state.  
We propose a approximate simulation method to study XXX


Continue a line of work by \citet{huang1995software} and 
\citet{dohi2000statistical}, this paper proposes an alternative Markov 
model for software rejuvenate.  Specifically, we expend each state, $S$, in 
traditional models to a group of states, $S(\tau)$, where $\tau$ represents 
the elapsed time that the system has been entered into state $S$.  
The above model is equivalent to a semi-Markov process.  It, however, brings two 
benefits and one limitation.  The first 
benefit is that state transition probabilities can be precisely described.  
Like in general reliability studies, the failure rate in the new model 
can be precisely described by an arbitrary probability density function or 
probability mass function.  In contrast, traditional models either use the 
average Mean Time to Failure (MTTF) for cost analysis or assume that the time 
spent in each state has an exponential distribution.  The second 
benefit is that transition probabilities of the new Markov process has take 
time into account.  The result model is either a simple Markov chain if time is 
a discrete value,  or a general Markov process if time is a continuous value.  
Comparing to traditional models that are semi-Markov processes or Continuous 
Time Markov Chains (CTMT), the Markov Chain used in proposed model is often 
easier to understand and analysis.  The downside of expending the state space 
is the decreased performance of simulation processes.  Fortunately, approximate 
estimations can be given to balance the trade-off between accuracy and 
efficiency.  Most importantly, we show that the classical model given by 
\citep{huang1995software} is the least accurate approximation.  Moreover, 
inspired by economical theories, a notion of {\it utility} function
is proposes paper to unify the cost analysis and the availability 
analysis, which are usually treated separately in previous studies.



\end{abstract}
\end{comment} 

% \begin{IEEEkeywords}
% key1; key2; key3; 

% \end{IEEEkeywords}

% \IEEEpeerreviewmaketitle


\input{Introduction}
\input{classicmodels}
\input{newmodel}
\input{example}




\section{Efficiency}

In Example HuangB \& HuangC.  The time is set to 10 minutes.  It takes 1 hour to
simulate the first year.  Not bad!

The matrix is sparse, space efficiency can be improved.


TODO: measure the complexity 


\section{TODO List}


\begin{itemize}
  \item Examples from Dohi
  \item Space Efficiency improvement
  \item ! compose sub-systems
\end{itemize}



\section{Conclusion}


% \section*{Acknowledgment}


% The authors would like to thank... more thanks here


\bibliographystyle{IEEEtranN}
\bibliography{rejuvenation}

% \appendix
%  \input{appendix_correct}

% \received{September 2008}{March 2009}

\end{document}


