\documentclass[coverpage]{inftechrep}

\usepackage{comment}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{url}

\usepackage{paralist}
\usepackage{hyperref}
\usepackage[all]{hypcap}
\usepackage{verbatim}
\usepackage{array}
\usepackage{float}
\usepackage{multirow}
% \usepackage{rotating}
\usepackage{multicol}
\usepackage{longtable}
\usepackage{changepage}

\usepackage{color}
\usepackage{listings}   
% "define" Scala
\lstdefinelanguage{scala}{
  morekeywords={abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

% A comment in a draft (shouldn't appear in the final version).
\newcommand{\mycomment}[1]{\(\spadesuit\){\bf #1 }\(\spadesuit\)}
% Comment this out in the draft
% \newcommand{\mycomment}[1]{}
\newcommand{\pmcomment}[1]{\comment{PM}{#1}}
\newcommand{\pwtcomment}[1]{\comment{PT}{#1}}
\newcommand{\rscomment}[1]{\comment{RS}{#1}}

% line break in table.  e.g.
% Foo bar & \specialcell{Foo\\bar} & Foo bar \\    % vertically centered
% Foo bar & \specialcell[t]{Foo\\bar} & Foo bar \\ % aligned with top rule
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

  

\begin{document}
\title{Statically Typed Reliability \\Type-parametrized Actors and Their
Supervision (v.0)
}
\author{Jiansen HE}
% \reportnumber{25}
% \reportyear{1999}
% \reportmonth{5}
\keywords{actor, type, supervision tree, nameserver}
% \published{Presented at the WITCH99 Conference, Debrecen, Hungary, May 1999.}
\institutes{
  Laboratory for Foundations of Computer Science}
\maketitle

\begin{abstract}
Building robust concurrent programs, especially in distributed settings, is
challenging as it often involves complex program structures and dynamic
behaviours.  The robustness of concurrent programs could be improved by
\begin{inparaenum}[(i)]
 \item using static type checking to prevent certain data misuses, or
 \item employing failure recovery mechanisms such as supervision tree.
\end{inparaenum}
An important open question is to what extent those approaches could be merged.

This paper presents the design, implementation and preliminary evaluation
results of the TAkka library, where types of actor related components are
statically typed and operations are type checked at the earliest possibility.
We introduce actors parametrized by the type of messages they expect to
receive. We show that it is straightforward to construct
supervision trees of type-parametrized actors. We minimize the number of system
messages that may be handled by general users by providing standard supervision
strategies.

We evaluate the TAkka library by re-implementing 19 examples from other
OTP-like libraries.  our results show that TAkka has small runtime and
code size overheads compared with Akka. Finally, TAkka programs have similar
scalability comparing
to their Akka equivalent.

\mycomment{rewrite the abstract to reflect the purpose of this technical report.}

\mycomment{Let's show how static and dynamic type checking improve actor
programming}

\end{abstract}

\input{intro}
\input{background}
\input{actor}
\input{implementation}
\input{evaluation}


\section{Conclusion}

Programs written in Actor Model should be easy to reason their intentions
\cite{Hewitt:1973}.  Unfortunately, actors in Erlang and Akka accept
dynamically typed messages and therefore have vague contracts with their outside
world.  The TAkka library introduces a type-parameter for actor-related class.
The additional type-parameter specifies the intention of that actor's solo
communication interface. More importantly, with the help of type-parametrized
actors, the implementation of an actor can be statically checked against the
intention of that actor.

In addition to eliminating programming bugs and type errors, programmers would
like to have reliable failure recovery mechanisms for unexpected run-time
errors.  Supervision tree \cite{OTP} is such a mechanism that has improved the
reliability of a large telecom system written in Erlang.  We are glad to see
that type-parametrized actors can form supervision trees in the same way as
regular actors.

Lastly, tests show that building type-parametrized actor on top of Akka actor
does not introduce significant overheads, respecting to program size,
efficiency, and scalability.  The above test result is encouraging in the sense
that a large amount of previous actor library implementation could be re-used.
We expect similar results could be obtained in other OTP-like libraries such as
the future version of CloudHaskell.


% \acks

% Acknowledgments



\bibliographystyle{plain}
\bibliography{Jiansen}

\appendix
%\include{api}
%%\include{plan}

\end{document}
